\documentclass[Afour,sageh,times]{sagej}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{fontspec}
\usepackage{hyperref}
\usepackage{url}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{adjustbox}
\newfontfamily\hebrewfont{Arial}
\newcommand{\texthebrew}[1]{{\hebrewfont #1}}
\setcitestyle{numbers,square}

\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny\color{gray},
}

\begin{document}

\runninghead{Goldberg et al.}

\title{Distant Supervision from Cataloging Workflows: Multi-Strategy MARC-to-Wikidata Conversion for Hebrew Manuscript Collections}

\author{Alexander Goldberg\affilnum{1}, Gila Prebor\affilnum{2}, Avshalom Elmalech\affilnum{3}}

\affiliation{\affilnum{1}Department of Information Science and Applied Artificial Intelligence, Bar-Ilan University, Ramat Gan, Israel\\
\affilnum{2}Department of Information Science and Applied Artificial Intelligence, Bar-Ilan University, Ramat Gan, Israel\\
\affilnum{3}Department of Information Science and Applied Artificial Intelligence, Bar-Ilan University, Ramat Gan, Israel}

\corrauth{Alexander Goldberg, Department of Information Science and Applied Artificial Intelligence, Bar-Ilan University, Ramat Gan, Israel.}
\email{golbera3@biu.ac.il}

\begin{abstract}
Transforming library catalogs into Linked Open Data is essential for connecting isolated cultural heritage collections to the Semantic Web, yet automated conversion from MARC~21 records to Wikidata remains a largely unsolved problem---particularly for non-Latin-script collections where existing tools achieve low field coverage. We present the MHM (Mapping Hebrew Manuscripts) Pipeline, a multi-strategy system for automated conversion of Hebrew manuscript MARC records to Wikidata, achieving 96\% date coverage, 100\% genre coverage, an average of 73.6 Wikidata statements per manuscript (including 4,162 exemplar-of links to automatically created work items), and 7.5 statements per person entity---reducing data loss from 26.4\% to approximately 0\%.

Our approach combines six complementary strategies: (A)~three domain-specific Named Entity Recognition models trained via distant supervision from MARC structured fields, achieving 85.7\%, 95.9\%, and 99.99\% F1 for person, provenance, and contents extraction respectively; (B)~rule-based pattern extraction for Hebrew dates, physical descriptions, and genre/subject term mapping to Wikidata QIDs; (C)~multi-authority linking through Mazal/NLI, VIAF, and KIMA gazetteers; (D)~VIAF cluster harvesting for cross-referencing GND, LCCN, ISNI, and BnF identifiers; (E)~property-level data type validation against the WikiProject Manuscripts Data Model; and (F)~a distant-supervision genre classifier trained on 25,421 quality-filtered records using whole-token Hebrew keyword matching and a 3-sentence context window as input — with DictaBERT warm-started from the NER checkpoint, top-2 layers fine-tuned, focal loss ($\gamma=2$), and a NOTA abstention class — achieving micro-F1 of 0.918 ($\pm$0.003, 5-fold CV; best fold 0.921) on 8 genre classes and raising P136 coverage from 69\% to approximately 85\%.

We evaluate our system end-to-end on 100 Hebrew manuscripts from the National Library of Israel, producing 670 Wikidata items (570 persons and 100 manuscripts) with zero failures and zero duplicate claims. We quantify data loss across the conversion pipeline, identifying five categories of information that cannot transfer to Wikidata's data model, and demonstrate that the dominant loss (4,169 unmapped work titles) is structural rather than technical. All code, models, and training data generation scripts are open-source.

\end{abstract}

\begin{keywords}
Linked Open Data; Wikidata; MARC~21; named entity recognition; distant supervision; cultural heritage; Hebrew manuscripts; authority linking
\end{keywords}

\maketitle


%% ====================================================================
\section{Introduction}
\label{sec:introduction}
%% ====================================================================

\subsection{The MARC-to-LOD Challenge}

Libraries worldwide hold vast collections of cultural heritage materials cataloged in MARC~21 (Machine-Readable Cataloging), a standard developed in the 1960s that remains the dominant bibliographic record format~\cite{loc-marc}. While MARC records encode rich metadata---titles, dates, person names, physical descriptions, provenance histories---this information is trapped in a format designed for card catalog automation rather than Semantic Web interoperability.

The National Library of Israel (NLI) maintains 123,621 Hebrew manuscript MARC records, representing one of the world's largest collections of Jewish cultural heritage materials. These records contain structured fields (fixed-length codes, controlled headings) alongside unstructured note fields (MARC~500, 505, 561) where catalogers recorded detailed descriptions in Hebrew free text. The structured-unstructured duality creates both a challenge and an opportunity: rule-based conversion handles only the structured portion, while the note fields---which often contain the richest information about authorship, provenance, and contents---require natural language understanding.

Wikidata~\cite{vrandecic2014wikidata}, the collaborative knowledge base of the Wikimedia ecosystem, has emerged as the preferred Linked Open Data (LOD) target for cultural heritage institutions~\cite{van2020wikidata}. The WikiProject Manuscripts\footnote{\url{https://www.wikidata.org/wiki/Wikidata:WikiProject_Manuscripts}} defines a community-driven data model for manuscript description. However, no existing system achieves more than approximately 50\% field coverage when converting non-English MARC records to Wikidata.

\subsection{Research Gap}

Existing MARC-to-LOD conversion approaches fall into three categories, each with significant limitations:

\begin{enumerate}[leftmargin=*]
\item \textbf{Rule-based converters} (BIBFRAME~\cite{mcallister2017bibframe}, LD4L~\cite{cornell-ld4l}) handle structured MARC fields but cannot extract information from free-text notes, leaving approximately 60\% of available metadata unmapped.
\item \textbf{NER-based extractors} for cultural heritage~\cite{ehrmann2023named} exist for English and European languages but not for Hebrew bibliographic descriptions, and none integrate with Wikidata's specific property model.
\item \textbf{LLM-based approaches} (GPT-4 prompting) show promise for zero-shot entity extraction but are expensive (\$0.03--0.06/record), non-reproducible, and cannot operate offline---a critical requirement for institutions handling sensitive materials.
\end{enumerate}

No published work combines NER with rule-based extraction, multi-authority linking, and LOD enrichment in a single end-to-end pipeline that produces validated Wikidata items. Furthermore, no study has quantified the data loss inherent in MARC-to-Wikidata conversion, identifying which information is structurally untransferable.

\subsection{Contributions}

This paper makes seven contributions:

\begin{enumerate}[leftmargin=*]
\item A novel \textbf{distant supervision methodology} that exploits the dual structure of MARC records---using structured fields as automatic labels for training NER models on free-text note fields---eliminating the need for manual annotation.
\item Three \textbf{domain-specific NER models} for Hebrew manuscript descriptions, achieving state-of-the-art performance on person (85.7\% F1), provenance (95.9\% F1), and contents (99.99\% F1) extraction.
\item A \textbf{MARC~500 sentence-level classifier} (Colophon~ML) with two independent sigmoid heads that jointly detect colophon and provenance sentences inside general notes, re-routing them into P1684 (inscription) and the Provenance NER pipeline respectively --- extending owner/date/collection coverage beyond MARC~561.
\item A \textbf{VIAF cluster harvesting} technique that leverages a single VIAF match to automatically cross-reference four additional international authority identifiers (GND, LCCN, ISNI, BnF), increasing person entity LOD connectivity.
\item The first \textbf{end-to-end quantitative evaluation} of MARC-to-Wikidata conversion measuring both coverage (statements per entity) and data loss (what cannot transfer and why).
\item A \textbf{five-category taxonomy of data loss} in MARC-to-Wikidata conversion, demonstrating that the dominant loss (4,169 unmapped work titles) is structural rather than technical.
\item A \textbf{complete open-source system}\footnote{\url{https://github.com/alexandergolbergwix/pipeline}} processing 123,621 records without manual annotation, with all training data generated automatically, packaged as a cross-platform PyQt6 desktop application with a three-axis Source~$\times$~Type~$\times$~Role filter UI for reviewing and correcting extractions before Wikidata upload.
\end{enumerate}


%% ====================================================================
\section{Related Work}
\label{sec:related}
%% ====================================================================

\subsection{MARC-to-Linked Data Conversion}

The Library of Congress BIBFRAME initiative~\cite{mcallister2017bibframe} defines an RDF vocabulary for bibliographic data, but BIBFRAME conversion tools (marc2bibframe2) are rule-based and limited to structured MARC fields. The LD4L (Linked Data for Libraries) project~\cite{cornell-ld4l} explored enriching library records with LOD links but focused on monograph catalogs rather than manuscripts. MARC2WIKI, a community tool for Wikidata, provides basic field mapping but achieves low coverage on non-English records due to limited date parsing and authority resolution.

The Digital Scriptorium\footnote{\url{https://digital-scriptorium.org/}} has pioneered manuscript description on Wikidata using the WikiProject Manuscripts data model, but their workflow requires substantial manual curation. Our system automates this process while maintaining compliance with the same data model.

\subsection{NER for Cultural Heritage}

Named Entity Recognition for historical documents has advanced significantly~\cite{ehrmann2023named}, with systems like HIPE~\cite{ehrmann2022extended} establishing benchmarks for multilingual historical NER. However, these focus on newspaper text rather than bibliographic descriptions, and none target Hebrew manuscript catalogs.

DictaBERT~\cite{shmidman2023dictabert}, a Hebrew BERT model pre-trained on diverse Hebrew corpora including historical texts, provides a strong foundation for domain-specific fine-tuning. Our prior work~\cite{goldberg2025person} demonstrated that DictaBERT can be fine-tuned for person NER on MARC note fields; this paper extends that approach to three entity types with a novel distant supervision methodology.

\subsection{Distant Supervision and Data Programming}

Distant supervision~\cite{mintz2009distant} generates training labels by aligning entity mentions in text with entries in a knowledge base. Snorkel~\cite{ratner2017snorkel} generalized this to data programming with labeling functions. Our approach differs from both: rather than using an external knowledge base, we exploit the \emph{internal dual structure} of MARC records where structured fields (100/700 for persons, 505 for contents) provide labels for the same information expressed as free text in note fields (500, 561).

\subsection{Wikidata for Cultural Heritage}

Wikidata has become a hub for cultural heritage LOD~\cite{van2020wikidata}, with initiatives like WikiProject~Books, WikiProject~Manuscripts, and Sum~of~All~Paintings linking millions of cultural objects. Feitosa et al.~\cite{feitosa2023wikidata} surveyed Wikidata usage in libraries, finding growing adoption but limited automation. Our work addresses this gap by providing a fully automated pipeline from MARC to Wikidata with property-level type validation.


%% ====================================================================
\section{System Architecture}
\label{sec:architecture}
%% ====================================================================

The MHM Pipeline consists of six stages, each producing an intermediate artifact that can be independently inspected and validated.

\subsection{Pipeline Overview}

\begin{enumerate}[leftmargin=*]
\item \textbf{MARC Parsing} (\texttt{UnifiedReader} + \texttt{field\_handlers.py}): Extracts 59 structured fields from MARC~21 records into a normalized \texttt{ExtractedData} representation, handling MARC's complex subfield structure and encoding variations.

\item \textbf{NER Extraction} (3 models): Applies three DictaBERT-based NER models in parallel to different MARC fields:
  \begin{itemize}
    \item Person NER on MARC~500 (general notes): extracts person names with role classification
    \item Provenance NER on MARC~561 (ownership history): extracts owners, dates, collections
    \item Contents NER on MARC~505 (contents): extracts work titles, folio references, work authors
  \end{itemize}

\item \textbf{Authority Resolution}: Matches extracted names against three authority databases: Mazal/NLI (local, 630K+ records), VIAF (international, SRU API), and KIMA (Hebrew place gazetteer). VIAF cluster harvesting extracts cross-referenced GND, LCCN, ISNI, and BnF identifiers.

\item \textbf{RDF Construction}: Maps enriched records to the Hebrew Manuscripts Ontology (HMO), producing an RDF graph compliant with FRBRoo and CIDOC-CRM.

\item \textbf{SHACL Validation}: Validates the RDF graph against SHACL shapes encoding cardinality constraints, value type requirements, and mandatory properties.

\item \textbf{Wikidata Upload}: Converts validated records to WikibaseIntegrator item representations and uploads via the Wikidata API with OAuth~2.0 authentication, claim-level diffing (no duplicate statements), and batch rate limiting.
\end{enumerate}

\subsection{Wikidata Property Mapping}

The pipeline maps to 37 Wikidata properties following the WikiProject Manuscripts data model. Table~\ref{tab:manuscript-props} shows manuscript properties; Table~\ref{tab:person-props} shows person properties.

\begin{table}[t]
\caption{Wikidata properties for manuscript entities.}
\label{tab:manuscript-props}
\centering
\small
\begin{tabular}{llll}
\toprule
\textbf{Property} & \textbf{Label} & \textbf{Source} & \textbf{Coverage} \\
\midrule
P31  & instance of      & Rule       & 100\% \\
P195 & collection       & Hardcoded  & 100\% \\
P17  & country          & Hardcoded  & 100\% \\
P131 & located in       & Hardcoded  & 100\% \\
P217 & inventory no.    & MARC 852   & 100\% \\
P8189& NLI J9U ID      & MARC 001   & 100\% \\
P1476& title            & MARC 245   & 99\%  \\
P407 & language         & MARC 008   & 100\% \\
P282 & writing system   & Rule       & 100\% \\
P571 & inception        & MARC 008   & 96\%  \\
P136 & genre            & MARC 655 + ML & 85\%  \\
P2635& number of parts  & CU count   & 99\%  \\
P921 & main subject     & MARC 630   & 94\%  \\
P7153& significant place& KIMA       & 82\%  \\
P1071& location created & KIMA       & 79\%  \\
P50  & author           & NER+Auth   & 100\% \\
P1684& inscription      & Colophon   & 41\%  \\
P11603&transcribed by   & NER+Auth   & 19\%  \\
P127 & owned by         & Prov.\ NER & 17\%  \\
P6216& copyright        & Rule       & 100\% \\
P5008& on focus list    & Hardcoded  & 100\% \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t]
\caption{Wikidata properties for person entities (verified on 570 persons).}
\label{tab:person-props}
\centering
\small
\begin{tabular}{lllr}
\toprule
\textbf{Property} & \textbf{Label} & \textbf{Strategy} & \textbf{Count} \\
\midrule
P31   & instance of    & Rule            & 570 \\
P106  & occupation     & Role mapping    & 314 \\
P214  & VIAF ID        & VIAF search     & 142 \\
P1343 & described by   & Hardcoded       & 570 \\
P21   & sex/gender     & Heuristic       & 332 \\
P1412 & language       & Hardcoded       & 332 \\
P1559 & native name    & Hebrew name     & 332 \\
P244  & LCCN           & Cluster harvest & 104 \\
P213  & ISNI           & Cluster harvest & 129 \\
P227  & GND ID         & Cluster harvest & 68  \\
P268  & BnF ID         & Cluster harvest & 29  \\
P569  & date of birth  & Mazal DB        & 12  \\
P570  & date of death  & Mazal DB        & 12  \\
\midrule
\multicolumn{3}{l}{\textbf{Average statements per person}} & \textbf{7.5} \\
\bottomrule
\end{tabular}
\end{table}


%% ====================================================================
\section{Multi-Strategy Extraction Methodology}
\label{sec:methodology}
%% ====================================================================

\subsection{Strategy A: Distant Supervision NER}
\label{sec:distant-supervision}

The core methodological contribution is the exploitation of MARC's dual structure for distant supervision. MARC records contain the same information twice: once in structured fields with controlled vocabulary (e.g., field~100\$a for author headings) and once in free-text notes written by catalogers (e.g., field~500\$a). This creates a natural alignment between structured labels and unstructured text without any manual annotation.

\subsubsection{Model 1: Person NER (85.7\% F1)}

\textbf{Training data generation.} For each MARC record, person names from structured fields (100\$a, 700\$a) are matched against occurrences in general note fields (500\$a) using exact and fuzzy string matching. When a name appears in the note text, the corresponding tokens are labeled as B-PERSON/I-PERSON; all other tokens receive the O label. This process generated 7,614 training samples from the NLI catalog.

\textbf{Architecture.} DictaBERT~\cite{shmidman2023dictabert} (110M parameters) with a token classification head. Trained with 5-fold stratified cross-validation on Apple~M1~Max, 2.5~hours total.

\textbf{Role classification.} The model extracts person mentions; role assignment (author, scribe, owner, translator, commentator, censor) uses keyword-based heuristics on surrounding Hebrew context, matching terms such as \texthebrew{מעתיק} (copyist), \texthebrew{בעלים} (owner), and \texthebrew{מתרגם} (translator).

\subsubsection{Model 2: Provenance NER (95.9\% F1)}

\textbf{Training data generation.} Three distant supervision strategies target MARC~561 (ownership/custodial history):
\begin{enumerate}[leftmargin=*]
\item \emph{Structured cross-reference}: Owner names from field~561 are aligned with authority headings in 100/700 fields of the same record.
\item \emph{Hebrew pattern matching}: Regular expressions identify ownership formulae (\texthebrew{ציון בעלים}, \texthebrew{אוסף}) and extract adjacent entity spans.
\item \emph{Latin censor identification}: Names following Latin censor attestation patterns (``Visto per mi...'') are labeled as PERSON entities.
\end{enumerate}

Training on 12,100 samples (28.4\% augmented with multi-entity examples) yielded 95.91\% F1 (best fold: 96.17\%) for three entity types: OWNER, DATE, COLLECTION.

\subsubsection{Model 3: Contents NER (99.99\% F1)}

MARC~505 (contents notes) follows a highly regular semi-structured format: numbered entries with work titles, folio references, and sometimes author names. Rule-based parsing of 505 fields generates training data with near-perfect labels, explaining the exceptional F1 score. Three entity types: WORK, FOLIO, WORK\_AUTHOR. Trained on 10,000 samples.

\subsection{Strategy B: Rule-Based Pattern Extraction}

\subsubsection{Hebrew Date Parsing}

Hebrew manuscript dates present unique challenges. NLI catalogs encode dates in three formats: structured years (``1495''), English century descriptions (``16th century''), and Hebrew century strings (\texthebrew{מאה ט"ז} = 16th century = ca.~1550). Our parser handles all three formats, including Hebrew ordinal-to-integer conversion via a mapping of 20 Hebrew letter-number combinations. Coverage improved from 22\% to 96\% after implementing Hebrew century parsing.

\subsubsection{Genre and Subject QID Mapping}

We created comprehensive mapping dictionaries: 50~genre terms to Wikidata QIDs (covering 100\% of NLI's genre vocabulary), 30~LCSH subject headings, 13~Bible books, and 14~Talmud Bavli tractates. All mappings are centralized in a single module (\texttt{property\_mapping.py}) for maintainability.

\subsection{Strategy C: Multi-Authority Linking}

Three authority databases provide entity resolution:

\textbf{Mazal/NLI} (local SQLite, 630K+ records): Fuzzy matching of person, place, work, and corporate body names against the NLI authority file. Provides NLI~J9U identifiers (P8189) and biographical dates.

\textbf{VIAF} (public API, SRU/CQL): Name search with JSON content negotiation. Rate-limited to 2~requests/second. Provides VIAF~URIs (P214).

\textbf{KIMA} (Hebrew place gazetteer, SQLite): Maps Hebrew historical place names to Wikidata QIDs via variant name matching. Provides location of creation claims (P1071).

\subsection{Strategy D: VIAF Cluster Harvesting}

When a person matches a VIAF cluster, we fetch the full cluster JSON from \texttt{https://viaf.org/viaf/\{id\}} (with \texttt{Accept: application/json} header) and extract cross-referenced authority identifiers. The VIAF cluster aggregates records from 40+ national libraries; a single match thus yields up to four additional external identifiers:

\begin{itemize}[leftmargin=*]
\item \textbf{P227} (GND): Deutsche Nationalbibliothek --- 68 persons (12\%)
\item \textbf{P244} (LCCN): Library of Congress --- 104 persons (18\%)
\item \textbf{P213} (ISNI): International Standard Name Identifier --- 129 persons (23\%)
\item \textbf{P268} (BnF): Biblioth\`{e}que nationale de France --- 29 persons (5\%)
\end{itemize}

This technique transforms a single VIAF URI into a web of cross-referenced international authority links, significantly increasing each person entity's LOD connectivity at minimal additional API cost (one HTTP request per person).

\subsection{Strategy E: Property-Level Validation}

Wikidata properties have strict value type constraints. Our validator checks every generated claim against the expected data type:
\begin{itemize}[leftmargin=*]
\item P8189 and P214 require \texttt{external-id}, not \texttt{string}
\item P5816 (condition) requires item QIDs, not free text --- claims with string values are skipped
\item P527 (has parts) requires item QIDs --- work titles without Wikidata items are omitted
\item P195 (collection) requires item QIDs --- NER-extracted collection names without reconciled QIDs are skipped
\end{itemize}

This validation prevents upload errors and ensures WikiProject Manuscripts Data Model compliance.

\subsection{Strategy F: Distant-Supervision Genre Classifier}

31\% of NLI Hebrew manuscript records have no MARC~655 genre/form headings, leaving P136 (genre) empty after rule-based QID mapping. To fill this gap we train a multi-label sequence classifier via distant supervision: records that \emph{do} carry MARC~655 labels provide (title + context window, genre) training pairs, and the trained model is applied to the 31\% without labels.

\textbf{Data and filtering.} The full Ktiv catalog export (123,621 records) is used as the training corpus, augmented with 897 intensively annotated records. A record is included only if three conditions hold: (1)~MARC~245 (title), MARC~500 (general notes), and MARC~655 (genre) are all present; (2)~at least one genre keyword appears in MARC~500 notes as a \emph{whole token}; and (3)~the genre maps to a key in \texttt{GENRE\_TO\_QID}. Condition~(2) is a quality gate: it confirms the cataloger explicitly described the genre in free text, ensuring a genuine textual signal exists for learning. Whole-token matching (space-boundary for Hebrew, \texttt{\textbackslash b} for ASCII) prevents spurious substring matches (e.g., \texthebrew{שיר} matching \texthebrew{שירות}). A pre-extraction step saves the 26,758 qualifying records to a compact TSV (\texttt{genre\_samples.tsv}), making all subsequent training runs fast without re-scanning the full catalog.

Two additional quality filters are applied: (i)~\emph{Literature (Miscellaneous, in manuscript)} is excluded because its keywords (\texthebrew{קובץ}, \texthebrew{אסופה}, ``collection'') are generic enough to appear across all genres, making its 2,170 samples predominantly false positives; (ii)~genre classes with fewer than 100 keyword-matched training examples are discarded (Business records: 75, Family records: 63, Biographies: 34, Negotiable instruments: 20), as they contribute more label noise than learning signal. After both filters, the training set contains 25,421 records across 8 high-confidence classes plus 1,629 NOTA examples.

\textbf{Keyword synonym dictionary.} Since MARC~655 headings are in English while MARC~500 notes are in Hebrew, direct string overlap yields near-zero matches. We therefore compiled \texttt{GENRE\_KEYWORDS}: a dictionary mapping each of the 20 genre classes to all known Hebrew morphological forms and English variants. For example, \emph{Legislation (Jewish law)} maps to \{תקנה, תקנות, תקנן, תוקן, חוק, חוקים, פסק, פסקי דין, גזרה, גזרות, חרם, חרמות, צו, צווים, takkanot, decree, ordinance, statute\}.

\textbf{Input (X).} Rather than using the entire notes field, the model input is the MARC~245 title concatenated with a \emph{3-sentence context window} from MARC~500: the sentence immediately before the keyword match, the matching sentence itself, and the sentence immediately after. This localises the input to the exact textual evidence the cataloger wrote about that genre. Sequences are truncated to 128 tokens.

\textbf{Architecture.} The encoder is \texttt{dicta-il/dictabert}~\cite{shmidman2023dictabert} warm-started from the provenance NER checkpoint (already domain-adapted on 12,100 Hebrew manuscript samples via our earlier distant supervision pipeline). The bottom 10 of 12 transformer layers are frozen — they encode generic Hebrew morphology and syntax already learned during NER fine-tuning — while the top 2 layers and the classification head are fine-tuned. This partial fine-tuning allows the encoder to adapt its higher-level representations to the genre domain without catastrophic forgetting of Hebrew pretraining. The [CLS] representation is passed through Dropout(0.3) and a linear layer to logits over 9 classes: the 8 retained genres plus an explicit \emph{None-of-the-Above} (NOTA) class, allowing the model to abstain when no genre applies. Focal loss~\cite{lin2017focal} with $\gamma=2$ and per-class positive weights $w_c = n_{\text{neg},c}/n_{\text{pos},c}$ is used instead of standard BCE, downweighting easy majority-class examples and focusing training on the harder, rarer classes.

\textbf{Training.} AdamW with differential learning rates: top 2 encoder layers at $2 \times 10^{-6}$ (10$\times$ lower than the head at $2 \times 10^{-5}$) to prevent catastrophic forgetting. Batch size 64, max sequence length 64 tokens. 5-fold stratified cross-validation, up to 30 epochs per fold, early stopping (patience~=~5), linear warmup over 10\% of steps. The best-fold checkpoint is saved. The classification threshold is tuned on each fold's validation set by scanning $[0.2, 0.8]$ in steps of 0.05 to maximise micro-F1; the optimal threshold is stored in the checkpoint and used at inference. Total training time on Apple M4 Pro (MPS): approximately 1.5 hours.

\textbf{Output and integration.} At inference, the classifier is applied to any record lacking MARC~655 genres. Because the model was trained on short 3-sentence context windows (typically under 64 tokens) but the inference input combines the title with up to three full MARC~500 notes (which can be substantially longer), a \emph{sliding window} strategy is used: the full token sequence is divided into overlapping 64-token windows with 50\% stride, each window is scored independently, and the sigmoid probabilities are averaged across all windows before thresholding. This prevents information loss from naive truncation. Genres predicted above the tuned threshold are added as P136 statements with epistemic qualifiers: P1480 = Q18122778 (\emph{presumably}) and a P887 reference citing Q2539 (\emph{machine learning}), distinguishing classifier-inferred claims from authority-sourced data. When the model predicts NOTA — or no genre exceeds the threshold — the classifier returns \emph{``other''} and no P136 claim is written. The model is loaded lazily and degrades gracefully when absent.

Table~\ref{tab:genre-classifier} summarises the 8 retained genre classes and their training set sizes.

\begin{table}[t]
\caption{Genre classifier training set: 8 retained classes (min.\ 100 examples) after quality filtering.}
\label{tab:genre-classifier}
\centering
\small
\begin{tabular}{lr}
\toprule
\textbf{Genre} & \textbf{N} \\
\midrule
Piyyutim & 5,852 \\
Poetry & 5,464 \\
Illustrated works (Manuscript) & 5,062 \\
Personal correspondence & 3,419 \\
Censored manuscripts & 2,364 \\
Autograph manuscripts & 2,122 \\
Records (Documents) & 1,052 \\
Bibliographies & 775 \\
\midrule
NOTA (abstention class) & 1,629 \\
\midrule
\textbf{Total} & \textbf{25,421} \\
\bottomrule
\end{tabular}
\end{table}

This strategy raises P136 coverage from 69\% to approximately 85\%.


%% ====================================================================
\section{Data Loss Analysis}
\label{sec:data-loss}
%% ====================================================================

A critical contribution of this work is the systematic quantification of data loss across the MARC-to-Wikidata conversion pipeline. Of 59~fields extracted from MARC records, 37~now map to Wikidata properties (up from 30 after targeted property research). We reduced the data loss rate from 26.4\% to approximately 17\% and identify five remaining categories of loss (Table~\ref{tab:data-loss}).

\begin{table*}[t]
\caption{Taxonomy of data loss in MARC-to-Wikidata conversion (100 manuscripts).}
\label{tab:data-loss}
\centering
\small
\begin{tabular}{p{3.5cm}p{4cm}rp{5.5cm}}
\toprule
\textbf{Category} & \textbf{Fields} & \textbf{Records} & \textbf{Explanation} \\
\midrule
\textbf{1. QID required,}\newline only text available
  & contents (work titles), condition\_notes, related\_works
  & 100, 2, 2
  & Wikidata properties P1574, P5816, P527 require item QIDs. 4,169 work titles have no Wikidata items. \\
\addlinespace
\textbf{2. No Wikidata}\newline property exists
  & has\_multiple\_hands
  & 3
  & No standard Wikidata property for multi-scribe detection. \\
\addlinespace
\textbf{3. Unstructured}\newline free text
  & notes
  & 100
  & General catalog notes with no structured target property. \\
\addlinespace
\textbf{4. Needs further}\newline entity resolution
  & 16 subject terms
  & 16
  & LCSH terms without QID mappings (person names misclassified as subjects). \\
\addlinespace
\textbf{5. Pipeline-internal}\newline metadata
  & attribution\_sources, hierarchy\_type, data\_from\_colophon
  & 100, 100, 41
  & Provenance tracking not intended for Wikidata. \\
\bottomrule
\end{tabular}
\end{table*}

\textbf{The dominant loss was Category~1}: 4,169~work titles across 100~manuscripts initially produced 0 claims because P1574 (exemplar~of) requires item QIDs. We resolved this by automatically creating Wikidata items for all 3,970~unique works (typed as Q47461344, with title, language, and author when available from NER). This eliminated the single largest data loss category and added 4,162~P1574 claims.

We reduced data loss from 26.4\% to approximately 17\% by identifying Wikidata properties for previously unmapped fields: P2635 (number of codicological units, 99~records), P1684 (colophon text as inscription with Q372474 qualifier, 41~records), P7153 (significant places via KIMA resolution, 82~claims), P3132 (explicit/last line, 2~records), and P478 (volume info, 1~record). Scribal interventions are mapped to P1684 with role qualifiers (Q860740 gloss, Q3299332 correction, Q1136474 marginalia).

Categories~2 and~3 represent \emph{irreducible structural loss}: general catalog notes have no structured Wikidata target, and multi-scribe detection lacks a standard property. Category~4 contains 16~remaining LCSH subject terms (mostly person names misclassified as subjects) awaiting QID mapping.


%% ====================================================================
\section{Experimental Evaluation}
\label{sec:evaluation}
%% ====================================================================

\subsection{Dataset}

We evaluate on 123,621~MARC records from the NLI Hebrew manuscript catalog. The test set comprises 100~manuscripts selected for maximum data richness: all contain MARC fields~500, 505, and~561, ensuring all three NER models can be evaluated. This selection strategy tests the pipeline at maximum capacity rather than on typical records.

\subsection{NER Performance}

Table~\ref{tab:ner-results} reports NER performance measured with 5-fold stratified cross-validation. All models use DictaBERT~\cite{shmidman2023dictabert} (110M parameters) as the base encoder.

\begin{table}[t]
\caption{Model performance (5-fold CV micro-F1).}
\label{tab:ner-results}
\centering
\small
\begin{tabular}{lrrrl}
\toprule
\textbf{Model} & \textbf{P} & \textbf{R} & \textbf{F1} & \textbf{Training} \\
\midrule
Person NER       & 86.0 & 85.4 & 85.7  & 2.5h (M1 Max)  \\
Provenance NER   & 94.0 & 96.0 & 95.9  & 40min (M4 Pro)  \\
Contents NER     & 100  & 100  & 99.99 & 15min (M4 Pro)  \\
Genre classifier & —    & —    & \textbf{0.918} $\pm$ 0.003  & $\sim$2h (M4 Pro) \\
\bottomrule
\end{tabular}
\end{table}

The provenance model's high F1 validates the distant supervision approach: 12,100~training samples generated automatically from MARC field alignment, with 28.4\% augmented with multi-entity examples, yielding a 1.95 percentage point improvement over single-entity training. The contents model's near-perfect F1 reflects the highly regular structure of MARC~505 fields. The genre classifier achieves micro-F1 of 0.918 $\pm$ 0.003 (best fold: 0.921) on 8 classes via a combination of quality-filtered distant supervision (whole-token Hebrew keyword matching, min.\ 100 examples per class), focal loss with per-class positive weighting, and partial fine-tuning of the top 2 DictaBERT layers warm-started from the provenance NER checkpoint. Table~\ref{tab:genre-perclass} reports per-class precision, recall, and F1 averaged across the 5 folds; the weakest class is \emph{Records (Documents)} (F1 0.796), reflecting its smaller training set (1,052 examples) and greater lexical diversity.

\begin{table}[t]
\caption{Genre classifier per-class results (mean across 5 folds, threshold = 0.65).}
\label{tab:genre-perclass}
\centering
\small
\begin{tabular}{lrrr}
\toprule
\textbf{Genre} & \textbf{P} & \textbf{R} & \textbf{F1} \\
\midrule
Autograph manuscripts      & 0.994 & 0.942 & 0.967 \\
Personal correspondence    & 0.972 & 0.950 & 0.961 \\
Piyyutim                   & 0.972 & 0.900 & 0.934 \\
Censored manuscripts       & 0.952 & 0.918 & 0.935 \\
Poetry                     & 0.971 & 0.875 & 0.920 \\
Illustrated works          & 0.982 & 0.847 & 0.910 \\
Bibliographies             & 0.908 & 0.854 & 0.880 \\
Records (Documents)        & 0.776 & 0.821 & 0.796 \\
\midrule
NOTA (abstention)          & 0.753 & 0.879 & 0.811 \\
\midrule
\textbf{Micro-F1}          &       &       & \textbf{0.918} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{End-to-End Coverage}

Table~\ref{tab:coverage} reports property-level coverage on the 100-manuscript test set.

\begin{table}[t]
\caption{Property-level coverage comparison.}
\label{tab:coverage}
\centering
\small
\begin{tabular}{lrrl}
\toprule
\textbf{Property} & \textbf{Rule} & \textbf{MHM} & \textbf{Strategy} \\
\midrule
P31 (type)       & 100\% & 100\%        & Rule \\
P1476 (title)    & 100\% & 100\%        & Rule \\
P17 (country)    & 0\%   & 100\%        & Hardcoded \\
P131 (located)   & 0\%   & 100\%        & Hardcoded \\
P571 (date)      & 22\%  & \textbf{96\%}& Hebrew parsing \\
P50 (author)     & 1/MS  & \textbf{7.6/MS}& NER + Authority \\
P136 (genre)     & 0\%   & \textbf{85\%}& QID mapping + ML \\
P2635 (parts)    & 0\%   & \textbf{99\%}& CU count \\
P921 (subject)   & 0\%   & \textbf{98\%}& Canonical+LCSH \\
P7153 (place)    & 0\%   & \textbf{82\%}& KIMA \\
P1071 (location) & 0\%   & \textbf{79\%}& KIMA \\
P1684 (colophon) & 0\%   & \textbf{41\%}& Text extraction \\
P127 (owner)     & 0\%   & \textbf{17\%}& Provenance NER \\
P11603 (scribe)  & 0\%   & \textbf{19\%}& Person NER \\
\midrule
\textbf{Avg stmts/MS} & \textbf{$\sim$8} & \textbf{73.6} & Combined \\
\bottomrule
\end{tabular}
\end{table}

The multi-strategy approach yields a 9.2$\times$ improvement over rule-only conversion (73.6 vs.\ 8.0 stmts/MS). Each strategy contributes coverage that the others cannot provide: NER unlocks free-text entity extraction (P50: 1$\to$7.6 authors/MS), rule-based parsing handles Hebrew-specific date formats (P571: 22\%$\to$96\%), and authority linking provides LOD identifiers.

\subsection{Upload Results}

A pilot upload of 670~items to Wikidata (Table~\ref{tab:upload}) validates the pipeline's production readiness.

\begin{table}[t]
\caption{Wikidata upload results (pilot: 100 manuscripts).}
\label{tab:upload}
\centering
\small
\begin{tabular}{lr}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Total items uploaded   & 4,653 \\
\quad Works            & 3,970 \\
\quad Persons          & 583 \\
\quad Manuscripts      & 100 \\
Total statements       & 7,359 (manuscripts) \\
Avg stmts/MS           & 73.6 \\
Data loss              & $\sim$0\% \\
Duplicate claims       & 0   \\
\bottomrule
\end{tabular}
\end{table}

The zero-failure rate is achieved through property-level type validation (Strategy~E) and claim-level diffing using WikibaseIntegrator's \texttt{MERGE\_REFS\_OR\_APPEND} mode, which adds new claims without duplicating or deleting existing ones. Re-running the upload on previously processed records correctly identifies 199~unchanged items, demonstrating idempotency.

\subsection{Ablation Study}

Table~\ref{tab:ablation} shows the contribution of each pipeline component.

\begin{table}[t]
\caption{Ablation study: contribution of each strategy.}
\label{tab:ablation}
\centering
\small
\begin{tabular}{lrr}
\toprule
\textbf{Configuration} & \textbf{Stmts/MS} & \textbf{Stmts/Person} \\
\midrule
Full pipeline (v2.0)      & 73.6 & 7.5 \\
$-$ Work items + notes    & 24.8 & 7.5 \\
$-$ Data loss fixes       & 22.9 & 7.5 \\
$-$ VIAF cluster harvest  & 73.6 & 5.5 \\
$-$ Person NER            & 66.7 & 7.5 \\
$-$ Hebrew date parsing   & 72.2 & 7.5 \\
$-$ Genre QID mapping     & 72.9 & 7.5 \\
Rule-only (no NER/LOD)    & 8.0  & 2.0 \\
\bottomrule
\end{tabular}
\end{table}

Work item creation is the largest single contributor ($+$48.8 stmts/MS), followed by person NER ($+$6.9) and data loss fixes ($+$1.9). VIAF cluster harvesting increases person entity richness by 36\% (5.5$\to$7.5 stmts/person). Without work items and notes, the pipeline still produces 24.8 core structural statements per manuscript.


%% ====================================================================
\section{Discussion}
\label{sec:discussion}
%% ====================================================================

\subsection{Why 100\% Coverage is Structurally Impossible}

Our data loss taxonomy (Section~\ref{sec:data-loss}) demonstrates that the gap between MARC and Wikidata is not merely a technical challenge but a structural one. Three fundamental barriers exist:

\begin{enumerate}[leftmargin=*]
\item \textbf{Entity gap}: Wikidata properties like P1574 (exemplar~of) and P527 (has~parts) require item QIDs. Of 3,968~unique Hebrew works referenced in our test set, only 12 have Wikidata items. Creating items for the remaining works is a community task beyond automated conversion.
\item \textbf{Property gap}: Manuscript-specific concepts (colophons, scribal interventions, codicological unit structure) have no Wikidata properties and may never receive them---they represent domain knowledge more suited to specialized ontologies.
\item \textbf{Model gap}: Wikidata's entity-centric model cannot represent relational structures inherent in codicological description, such as ``folios~1--50 contain Work~A in Hand~1, folios~51--100 contain Work~B in Hand~2.''
\end{enumerate}

These barriers suggest that a dual-publication strategy---detailed RDF using domain ontologies (e.g., FRBRoo/HMO) alongside summary Wikidata items---is the appropriate approach for cultural heritage LOD.

\subsection{The Distant Supervision Advantage}

Our distant supervision methodology has a key practical advantage: it requires zero manual annotation. The MARC dual structure---structured fields as labels, note fields as training text---is an inherent property of cataloging workflows worldwide. Any institution with MARC records containing both controlled headings and descriptive notes can replicate our approach for their collection, regardless of language or subject domain.

The methodology's effectiveness varies with cataloging quality: NLI's records, created by professional catalogers over decades, provide high-quality distant labels. Less rigorously cataloged collections may require additional noise handling.

\subsection{VIAF Cluster Harvesting as LOD Amplification}

The cluster harvesting technique demonstrates that LOD enrichment is multiplicative: a single successful authority match (VIAF) cascades into 2--4 additional cross-references (GND, LCCN, ISNI, BnF), each connecting the entity to a different national library's knowledge graph. For the 142~persons with VIAF matches in our test set, cluster harvesting added 330~additional authority identifiers---an amplification factor of 2.3$\times$.

This technique is generalizable to any domain where VIAF covers the entities of interest. The rate-limiting constraint (2~requests/second for search + 1 additional for cluster fetch) means processing 123K~records requires approximately 17~hours for the authority resolution stage, which is acceptable for batch processing.

\subsection{Domain Expert Validation}

The pipeline's Wikidata output was reviewed by two domain experts in Hebrew manuscript studies from the University of Haifa (M.~Lavee and E.~Baumgarten, October 2025). Their feedback, grounded in hands-on cataloging of seven test manuscripts (Vatican~44, Huntington~115, Oppenheimer~129, Parma~3122, two London manuscripts, and Jerusalem~8210), identified four gaps in the Wikidata representation:

\begin{enumerate}[leftmargin=*]
\item \textbf{Certainty/confidence mechanism}: No formal way to express that a scribe attribution is ``probable'' vs.\ ``certain''. The experts requested a vocabulary equivalent to \emph{certain / probable / possible} on individual claims.
\item \textbf{Partial work identification}: Manuscripts frequently contain compositions that partially overlap or derive from known works without being identical to them. The experts requested relations analogous to \texttt{is\_variant\_of} or \texttt{possibly\_realises} (IFLA LRM terminology) to express this grey area.
\item \textbf{Multi-volume and anthology structure}: The BU--CU--PU hierarchy is insufficiently expressive for manuscripts where one codicological unit contains nested sub-units, or where multiple volumes collectively form a single work.
\item \textbf{Textual overlap between scribal hands}: No formal relation for cases like Huntington~115, where two scribes copied overlapping sections of the same text.
\end{enumerate}

We addressed items (1) and (2) directly: unconfirmed person attributions (those without an authority-matched Wikidata QID) now receive a \texttt{P1480~=~Q18122778} (\emph{presumably}) qualifier; unreconciled work links (P1574) receive the same qualifier, implementing the \texttt{possibly\_realises} semantics within Wikidata's property model. Items (3) and (4) remain structural limitations of Wikidata (see below) and are noted as directions for a companion HMO-based RDF publication.

\subsection{Community Validation and Safety Guards}

A pilot upload of 670~items to Wikidata in April~2026 generated substantive community feedback through user talk pages. The Wikidata community identified: (a)~duplicate person items caused by the reconciler checking only VIAF and NLI identifiers, missing existing items findable via LCCN/GND/ISNI; (b)~institutional holders (MARC~710 entries) incorrectly attached as P50~(author) instead of P195~(collection) or P127~(owned~by); (c)~P8189 authority-record identifiers attached to non-human entities; and (d)~generic catalog placeholders (Hebrew \textit{kovetz}, `collection') emitted verbatim as Wikidata labels.

Each issue was fixed and protected by a regression test. The full reconciler now checks all five identifier types before creating any new person entity; the institutional name detector routes MARC~710 entries to the appropriate collection/ownership property; P8189 is gated on both authority-record prefix and \texttt{P31=Q5} (human); and placeholder titles are detected and replaced with shelfmark-based labels. Eleven pre-upload validation checks in \texttt{item\_validator.py} block any item carrying a known failure pattern (empty label, anonymous person, kovetz placeholder, trailing punctuation, MARC-inverted label, institutional-as-person, P3959 on human, P8189 bad prefix, VIAF on non-person, multiple same-language P1559, no-identifier person) from being approved in the GUI; every error-level finding disables the ``Approved'' checkbox for that row.

Beyond data-quality checks, a secondary class of incident arose when an automated cleanup script modified items the pipeline had never created. To make this class of error structurally impossible, the uploader now enforces a \textbf{four-stage modification guard} (\texttt{CLAUDE.md} rule~38) gated by \textbf{three independent verification channels}. The four gates are: (1)~an entry-point check in \code{upload\_item()}; (2)~a re-assertion inside \code{\_build\_wbi\_item()} so new call paths cannot bypass gate~1; (3)~a per-statement identity-conflict check refusing any value that would create a multi-value conflict on an existing identifier; and (4)~a last-ditch check placed immediately before the single \code{wbi\_item.write()} call site in the code-base. Raising \code{UnauthorisedModificationError} at any gate converts the attempt into a clean \emph{skipped} result rather than a partial edit. The three verification channels, all cross-checked inside \code{\_is\_our\_item()}: (i)~the MediaWiki API's \code{prop=revisions\&rvdir=newer} endpoint for the authoritative first-revision author; (ii)~the MediaWiki API's \code{list=usercontribs\&uctype=new} endpoint as an independent confirmation that the authenticated user has a \emph{page-creation} edit on the QID; and (iii)~a Wikidata SPARQL \code{ASK} query confirming the QID still exists and has not been deleted, redirected, or blanked since reconciliation. A negative answer from any channel refuses the edit; a network failure on one channel does not unlock the gate as long as the primary revisions answer still agrees.

The 232~unit tests in \texttt{test\_safety\_guards.py} serve as a regression barrier ensuring these classes of error cannot recur. Three of those tests are \emph{structural} rather than behavioural --- they fail if a future refactor introduces a second \code{wbi\_item.write()} call, separates the pre-write guard from the write, or re-introduces a marker-based creator fallback --- so the architecture of Rule~38 cannot be unintentionally weakened.

\subsection{Limitations}

\begin{enumerate}[leftmargin=*]
\item \textbf{Test set bias}: Our 100-manuscript test set was selected for data richness. Typical NLI records contain fewer note fields, so average coverage on the full catalog will be lower.
\item \textbf{Person NER role classification}: The model's role classification head degenerates to always predicting AUTHOR. We use keyword heuristics as a workaround, but a dedicated role classification model would be more robust.
\item \textbf{VIAF rate limiting}: At 2~requests/second, full authority resolution of 123K~records requires significant processing time. Caching and batch scheduling mitigate but do not eliminate this constraint.
\item \textbf{Language specificity}: The Hebrew date parser, NER models, and keyword heuristics are Hebrew-specific. Adapting to other languages requires retraining NER models and creating new parsing rules, though the distant supervision methodology transfers directly.
\item \textbf{Wikidata codicological expressivity}: Manuscript-specific relational structures (nested scribal layers, textual overlap, variant-work relationships at sub-folio granularity) exceed what Wikidata's property model can represent. A dual-publication strategy pairing Wikidata summary items with a full HMO-based RDF graph is the planned long-term solution.
\end{enumerate}


%% ====================================================================
\section{Conclusion}
\label{sec:conclusion}
%% ====================================================================

We presented the MHM Pipeline, a multi-strategy system for automated MARC-to-Wikidata conversion that achieves 73.6~statements per manuscript and 7.5~statements per person entity on Hebrew manuscript records from the National Library of Israel, mapping to 37~Wikidata properties with approximately 0\% data loss. Automatic work item creation (3,970~items) eliminates the dominant structural loss identified in our taxonomy. Our five-strategy approach---distant supervision NER, rule-based extraction, multi-authority linking, VIAF cluster harvesting, and property-level validation---demonstrates that no single technique suffices for high-coverage bibliographic LOD conversion.

The key methodological contribution is the distant supervision paradigm exploiting MARC's dual structure: structured fields provide automatic labels for training NER models on free-text notes, eliminating manual annotation. This approach is replicable to any MARC catalog with both controlled headings and descriptive notes.

Our data loss taxonomy identifies five categories of information that cannot transfer to Wikidata, demonstrating that the dominant loss (4,169~unmapped work titles) is structural rather than technical. This finding has implications for the cultural heritage LOD community: comprehensive digital representation requires both summary Wikidata items for discoverability and detailed domain-specific RDF for scholarly use.

A verified pilot upload of 670~items to Wikidata achieved zero failures and zero duplicate claims, demonstrating production readiness. All code, models, and training data generation scripts are available under GPL-3.0 at \url{https://github.com/alexandergolbergwix/pipeline}.

\subsection{Future Work}

Five directions are planned: (1)~creating Wikidata items for the 3,968 unique Hebrew works identified in our contents analysis, enabling confirmed P1574 claims; (2)~evaluating the pipeline on the full 123,621-record NLI catalog; (3)~adapting the distant supervision methodology to Arabic and Latin manuscript collections to test cross-lingual transferability; (4)~replacing the \emph{presumably} qualifier on NER-identified works with confirmed Wikidata reconciliation once the work-item corpus is established; and (5)~publishing a companion HMO-based RDF graph to address the codicological expressivity gaps identified by domain experts---in particular, nested CU structure, textual overlap between scribal hands, and the \texttt{is\_variant\_of} relationship for grey-area work identification.


%% ====================================================================
%% REFERENCES
%% ====================================================================

\begin{thebibliography}{99}

\bibitem{loc-marc}
Library of Congress.
\newblock MARC~21 Format for Bibliographic Data.
\newblock \url{https://www.loc.gov/marc/bibliographic/}, 2024.

\bibitem{vrandecic2014wikidata}
Vrande\v{c}i\'{c}, D. and Kr\"{o}tzsch, M.
\newblock Wikidata: A Free Collaborative Knowledge Base.
\newblock \emph{Communications of the ACM}, 57(10):78--85, 2014.

\bibitem{van2020wikidata}
van Veen, T.
\newblock Wikidata: From ``an'' Identifier to ``the'' Identifier.
\newblock \emph{Information Technology and Libraries}, 38(2):72--81, 2020.

\bibitem{mcallister2017bibframe}
McCallum, S.
\newblock BIBFRAME Development.
\newblock \emph{JLIS.it}, 8(3):71--85, 2017.

\bibitem{cornell-ld4l}
Cornell University Library.
\newblock Linked Data for Libraries (LD4L).
\newblock \url{https://ld4l.org/}, 2016.

\bibitem{ehrmann2023named}
Ehrmann, M., Hamdi, A., Pontes, E.L., Romanello, M., and Doucet, A.
\newblock Named Entity Recognition and Classification in Historical Documents: A Survey.
\newblock \emph{ACM Computing Surveys}, 56(2):1--47, 2023.

\bibitem{ehrmann2022extended}
Ehrmann, M., Romanello, M., Fl\"{u}ckiger, A., and Clematide, S.
\newblock Extended Overview of HIPE-2022: Named Entity Recognition and Linking in Multilingual Historical Documents.
\newblock In \emph{CLEF 2022}, pp.~1038--1066, 2022.

\bibitem{shmidman2023dictabert}
Shmidman, A., Shmidman, S., Koppel, M., and Goldberg, Y.
\newblock DictaBERT: A State-of-the-Art BERT Suite for Modern Hebrew.
\newblock \emph{arXiv preprint arXiv:2304.09919}, 2023.

\bibitem{goldberg2025person}
Goldberg, A., Prebor, G., and Elmalech, A.
\newblock Person NER for Hebrew Manuscript Catalogs via Distant Supervision.
\newblock \emph{Submitted}, 2025.

\bibitem{mintz2009distant}
Mintz, M., Bills, S., Snow, R., and Jurafsky, D.
\newblock Distant Supervision for Relation Extraction without Labeled Data.
\newblock In \emph{ACL-IJCNLP 2009}, pp.~1003--1011, 2009.

\bibitem{ratner2017snorkel}
Ratner, A., Bach, S.H., Ehrenberg, H., Fries, J., Wu, S., and R\'{e}, C.
\newblock Snorkel: Rapid Training Data Creation with Weak Supervision.
\newblock In \emph{VLDB}, 11(3):269--282, 2017.

\bibitem{feitosa2023wikidata}
Feitosa, D., Dermaku, G., and Ioannidis, L.
\newblock Wikidata for Digital Libraries: A Systematic Literature Review.
\newblock \emph{Journal of Documentation}, 79(7):1--25, 2023.

\end{thebibliography}

\end{document}
